---
title: "Spark Pipelines"
output:
  md_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(nycflights13)
library(sparklyr)
library(dplyr)
```


Spark's **ML Pipelines** provide a way to easily combine multiple transformations and algorithms into a single workflow, or pipeline.  

For R users, the insights gathered during the interactive sessions with Spark can now be converted to a formal pipeline.  This makes the hand-off from Data Scientists to Big Data Engineers a lot easier, this is because there should not be additional changes needed to be made by the later group.

The final list of selected variables, data manipulation, feature transformations and modeling can be easily re-written into a `ml_pipeline()` object, saved, and ultimately placed into a Production environment.  The `sparklyr` output of a saved Spark ML Pipeline object is in Scala code, which means that the code can be added to the scheduled Spark ML jobs, and without any dependencies in R.

## Introduction to ML Pipelines

The official Apache Spark site contains a more complete overview of [ML Pipelines](http://spark.apache.org/docs/latest/ml-pipeline.html).  This article will focus in introducing the basic concepts and steps to work with ML Pipelines via `sparklyr`.

There are two important stages in building an ML Pipeline.  The first one is creating a **Pipeline**.  A good way to look at it, or call it, is as an **"empty" pipeline**.  This step just builds the steps that the data will go through. This is the somewhat equivalent of doing this in R:

```{r}
r_pipeline <-  . %>% mutate(cyl = paste0("c", cyl)) %>% lm(am ~ cyl + mpg, data = .)
r_pipeline
```

The `r_pipeline` object has all the steps needed to transform and fit the model, but it has not yet transformed any data. 

The second step, is to pass data through the pipeline, which in turn will output a fitted model. That is called a **PipelineModel**.  The **PipelineModel** can then be used to produce predictions.  


```{r}
r_model <- r_pipeline(mtcars)
r_model
```


### Taking advantage of Pipelines and PipelineModels

The two stage ML Pipeline approach produces two final data products:

- A **PipelineModel** that can be added to the daily Spark jobs which will produce new predictions for the incoming data, and again, with no R dependencies.

- A **Pipeline** that can be **easily re-fitted** on a regular interval, say every month.  All that is needed is to pass a new sample to obtain the new coefficients.


## Pipeline

An additional goal of this article is that the reader can follow along, so the data, transformations and Spark connection in this example will be kept as easy to reproduce as possible.

```{r}
library(nycflights13)
library(sparklyr)
library(dplyr)
sc <- spark_connect(master = "local", spark_version = "2.2.0")
spark_flights <- sdf_copy_to(sc, flights)
```

### Feature Transformers

Pipelines make heavy use of [Feature Transformers](http://spark.rstudio.com/reference/#section-spark-feature-transformers).  If new to Spark, and `sparklyr`, it would be good to review what these transformers do.  These functions use the Spark API directly to transform the data, and may be faster at making the data manipulations that a `dplyr` (SQL) transformation.  

In `sparklyr` the `ft` functions are essentially are wrappers to original [Spark feature transformer](http://spark.apache.org/docs/latest/ml-features.html).

### ft_dplyr_transformer

This example will start with `dplyr` transformations, which are ultimately SQL transformations, loaded into the `df` variable. 

In `sparklyr`, there is one feature transformer that is not available in Spark, `ft_dplyr_transformer()`.  The goal of this function is to convert the `dplyr` code to a SQL Feature Transformer that can then be used in a Pipeline.

```{r}
df <- spark_flights %>%
  filter(!is.na(dep_delay)) %>%
  mutate(
    month = paste0("m", month),
    day = paste0("d", day)
  ) %>%
  select(dep_delay, sched_dep_time, month, day, distance) 
```

This is the resulting pipeline stage produced from the `dplyr` code:
```{r, eval = FALSE}
ft_dplyr_transformer(sc, df)
```

Use the `ml_param()` function to extract the "statement" attribute. That attribute contains the finalized SQL statement.  Notice that the `flights` table name has been replace with `__THIS__`.  This allows the pipeline to accept different table names as its source, making the pipeline very modular. 

```{r}
ft_dplyr_transformer(sc, df) %>%
  ml_param("statement")
```

### Ceating the Pipeline

The following step will create a 5 stage pipeline:

1. SQL transformer - Resulting from the `ft_dplyr_transformer()` transformation
2. Binarizer - To determine if the flight should be considered delay.  The eventual outcome variable.
3. Bucketizer - To split the day into specific hour buckets
4. R Formula - To define the model's formula
5. Logistic Model 

```{r}
flights_pipeline <- ml_pipeline(sc) %>%
  ft_dplyr_transformer(
    tbl = df
    ) %>%
  ft_binarizer(
    input.col = "dep_delay",
    output.col = "delayed",
    threshold = 15
  ) %>%
  ft_bucketizer(
    input.col = "sched_dep_time",
    output.col = "hours",
    splits = c(400, 800, 1200, 1600, 2000, 2400)
  )  %>%
  ft_r_formula(delayed ~ month + day + hours + distance) %>% 
  ml_logistic_regression()
```

Another nice feature for ML Pipelines in `sparklyr`, is the print-out.  It makes it really easy to how each stage is setup:
```{r}
flights_pipeline
```

Notice that there are no *coefficients* defined yet. That's because no data has been actually processed. Even though `df` uses `spark_flights()`, recall that the final SQL transformer makes that name, so there's no data to process yet.


## PipelineModel

A quick partition of the data is created for this exercise.
```{r}
partitioned_flights <- sdf_partition(
  spark_flights,
  training = 0.01,
  testing = 0.01,
  rest = 0.98
)
```

The `ml_fit()` function produces the PipelineModel.  The `training` partition of the `partitioned_flights` data is used to train the model:
```{r}
fitted_pipeline <- ml_fit(
  flights_pipeline,
  partitioned_flights$training
)
fitted_pipeline
```

Notice that the print-out for the fitted pipeline now displays the model's coefficients.

The `ml_transform()` function can be used to run predictions, in other words it is used instead of `predict()` or `sdf_predict()`.

```{r}
predictions <- ml_transform(
  fitted_pipeline,
  partitioned_flights$testing
)

predictions %>%
  group_by(delayed, prediction) %>%
  tally()
```

## Save the pipelines to disk

The `ml_save()` command can be used to save the Pipeline and PipelineModel to disk.  The resulting output is a folder with the selected name, which contains all of the necessary Scala scripts:
```{r, message = FALSE}
ml_save(
  flights_pipeline,
  "flights_pipeline",
  overwrite = TRUE
)

ml_save(
  fitted_pipeline,
  "flights_model",
  overwrite = TRUE
)
```

## Use an existing PipelineModel

The `ml_load()` command can be used to re-load Pipelines and PipelineModels.  The saved ML Pipeline files can only be loaded into an open Spark session.

```{r}
reloaded_model <- ml_load(sc, "flights_model")
```

A simple query can be used as the table that will be used to make the new predictions.  This of course, does not have to done in R, at this time the "flights_model" can be loaded into an independent Spark session outside of R.

```{r}
new_df <- spark_flights %>%
  filter(
    month == 7,
    day == 5
  )

ml_transform(reloaded_model, new_df) 
```

## Re-fit an existing Pipeline

First, reload the pipeline into an open Spark session:
```{r}
reloaded_pipeline <- ml_load(sc, "flights_pipeline")
```

Use `ml_fit()` again to pass new data, in this case, `sample_frac()` is used instead of `sdf_partition()` to provide the new data.  The idea being that the re-fitting would happen at a later date than when the model was initially fitted.
```{r}
new_model <-  ml_fit(reloaded_pipeline, sample_frac(spark_flights, 0.01))

new_model
```

The new model can be saved using `ml_save()`.  A new name is used in this case, but the same name as the existing PipelineModel to replace it.
```{r}
ml_save(new_model, "new_flights_model", overwrite = TRUE)
```

Finally, this example is complete by closing the Spark session.
```{r}
spark_disconnect(sc)
```





