---
title: "RSpark ML: Examples"
output:
  github_document:
    fig_width: 9
    fig_height: 5
---

## Initialization

```{r}
library(rspark)
library(dplyr)
library(ggplot2)

sc <- spark_connect("local", cores = "auto", version = "2.0.0-preview")
db <- src_spark(sc)

copy_to(db, iris, "iris")
iris_tbl <- tbl(db, "iris")
```

## KMeans in R

```{r}
cl <- iris %>%
  select(Petal.Width, Petal.Length) %>%
  kmeans(centers = 3)

centers <- as.data.frame(cl$centers)

iris %>%
  select(Petal.Width, Petal.Length) %>%
  ggplot(aes(Petal.Length, Petal.Width)) +
    geom_point(data = centers, aes(Petal.Width, Petal.Length), size = 60, alpha = 0.1) +
    geom_point(data = iris, aes(Petal.Width, Petal.Length), size = 2, alpha = 0.5)
```

## KMeans in RSpark

Basing kmeans over Spark on [spark.mllib K-means](http://spark.apache.org/docs/latest/mllib-clustering.html#k-means)

Note that the names of variables within the iris `tbl` have been transformed
(replacing `.` with `_`) to work around an issue in the Spark 2.0.0-preview
used in constructing this document -- we expect the issue to be resolved with
the release of Spark 2.0.0.

```{r}
model <- tbl(db, "iris") %>%
  select(Petal_Width, Petal_Length) %>%
  ml_kmeans(centers = 3)

tbl(db, "iris") %>%
  select(Petal_Width, Petal_Length) %>%
  collect %>%
  ggplot(aes(Petal_Length, Petal_Width)) +
    geom_point(data = model$centers, aes(Petal_Width, Petal_Length), size = 60, alpha = 0.1) +
    geom_point(aes(Petal_Width, Petal_Length), size = 2, alpha = 0.5)

```

## Linear Regression in R

```{r}
data <- iris %>%
  select(Petal.Width, Petal.Length)

model <- lm(Petal.Length ~ Petal.Width, data = iris)

iris %>%
  select(Petal.Width, Petal.Length) %>%
  ggplot(aes(Petal.Length, Petal.Width)) +
    geom_point(data = iris, aes(Petal.Width, Petal.Length), size = 2, alpha = 0.5) +
    geom_abline(aes(slope = coef(model)[["Petal.Width"]],
                    intercept = coef(model)[["(Intercept)"]],
                    color = "red"))
```

## Linear Regression in RSpark

```{r}
model <- tbl(db, "iris") %>%
  select(Petal_Width, Petal_Length) %>%
  ml_linear_regression(response = "Petal_Length", features = c("Petal_Width"))

iris %>%
  select(Petal.Width, Petal.Length) %>%
  ggplot(aes(Petal.Length, Petal.Width)) +
    geom_point(data = iris, aes(Petal.Width, Petal.Length), size = 2, alpha = 0.5) +
    geom_abline(aes(slope = coef(model)[["Petal_Width"]],
                    intercept = coef(model)[["(Intercept)"]],
                    color = "red"))
```

## Principal Components Analysis in R

```{r}
model <- iris %>%
  select(-Species) %>%
  prcomp()
print(model)

# calculate explained variance
model$sdev^2 / sum(model$sdev^2)
```

## Principal Components Analysis in RSpark

```{r}
model <- tbl(db, "iris") %>%
  select(-Species) %>%
  ml_pca()
print(model)
```

## Cleanup

```{r}
spark_disconnect(sc)
```
